---
title: Tests of the individual mean response have more power than tests of the area
  under curve (AUC) in glucose tolerance tests
author: "Jeffrey A. Walker"
date: "1/11/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(here)
library(ggplot2)
library(ggpubr)
library(cowplot)
library(data.table)
library(kableExtra)
```

# Methods
1. is a more powerful test. That is, it takes fewer samples to provide sufficient evidence of the direction of the treatment effect.
2. has the expected Type I error rate, while that of the comparison of individual time points is inflated.
3. has the expected Type I error rate conditional on the difference in baseline between treatments, while that of the comparison of AUC increases with the magnitude of the baseline difference.

# methods - make this into a table

```{r make methods table, echo=FALSE}
method_dt <- data.table(Methods = c("AUC",
                                    "IAUC",
                                    "Area-cov",
                                    "Mean",
                                    "Change",
                                    "Mean-cov",
                                    "Multi-t",
                                    "RM-ANOVA",
                                    "CLDA",
                                    "LDA-cov",
                                    "Roast",
                                    "O'Brien"),
                        Description = c("Area under the curve",
                  "Incremental area under the curve",
                  "Area under the post-baseline curve with baseline covariate",
                  "Mean of the post-baseline measures",
                  "Change from baseline of the post-baseline measures",
                  "Mean of the post-baseline measures with baseline covariate",
                  "Multiple t-tests",
                  "Repeated measures ANOVA",
                  "Constrained longitudinal data analysis",
                  "Longitudinal data analysis with baseline covariate",
                  "Rotational test of multivariate response",
                  "Non-parametric test of multivariate response"))
method_kable <- kable(method_dt, "latex", caption = "Methods for comparing the response between treatments", booktabs = T) %>%
  kable_styling() %>%
  pack_rows("Linear models of AUC", 1, 3) %>%
  pack_rows("Linear models of the mean response", 4, 6)  %>%
  pack_rows("Multivariate linear models", 7, 10)

method_kable
```

# Results
# Discussion
# References

