---
title: "Empirical examples of analysis of glucose tolerance tests"
author: "Jeff Walker"
date: "1/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(readxl)
library(janitor)
library(data.table)
library(ggpubr)
library(nlme)
library(mediation)
library(rms)
data_path <- "data"
```


```{r functions}
auc <- function(x, y, method="auc"){
  # method = "auc", auc computed using trapezoidal calc
  # method = "iauc" is an incremental AUC of Le Floch
  # method = "pos.iauc" is a "positive" incremental AUC of Le Floch but not Wolever
  # method = "pb.auc" is AUC of post-time0 values
  if(method=="iauc"){y <- y - y[1]}
  if(method=="pos.iauc"){y[y < 0] <- 0}
  if(method=="pb.auc"){
    x <- x[-1]
    y <- y[-1]
  }
  n <- length(x)
  area <- 0
  for(i in 2:n){
    area <- area + (x[i] - x[i-1])*(y[i-1] + y[i])
  }
  area/2
}

plot_all_curves <- function(data){
  gg <- ggplot(data=data, 
         aes(x = time,
             y = glucose,
             color = treatment)) +
    geom_point() +
    geom_line(aes(group = id))
 return(gg)
}

```
# Summary
## Sigma
## sigma
lead    0.2527439 0.5507195 0.5451951 0.4125610 0.6009756
ardb2   0.06581794 0.39155361 0.25379231 0.28717465
arnt    0.12818475 0.08265306 0.30683136 0.12818475 0.08265306 0.30683136
pi3k    0.4574595 0.8412703 1.6563243 2.3108919 1.9452703 2.0679189 1.6714054
march1  0.09597674 0.31565550 0.18280642 0.27442647 0.24691554
nrk1    0.06818708 0.06818708 0.06818708 0.06818708 0.06818708 0.06818708 0.06818708
beta    0.04268047 0.18612678 0.04268047 0.18612678 0.04268047 0.18612678
bcaa    0.1276123 0.2923751 0.2841936 0.1276123 0.2923751 0.2841936
il36    0.04192377 0.04192377 0.04192377 0.04192377 0.04192377

# Working template

```{r import-xxx, echo=FALSE, eval=FALSE}
folder <- "Recapitulating endocrine cell clustering in culture promotes maturation of human stem-cell-derived β cells"
filename <- "41556_2018_271_MOESM8_ESM.xlsx"
file_path <- here(data_path, folder, filename)
times <- c(0, 15, 30, 60, 90, 120)
time_cols <- paste0("glucose", times)

xxx_wide <- data.table(treatment = "eBCs",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "FIGURE 7",
                    range = "B5:F10")))
xxx_wide <- rbind(xxx_wide,
                    data.table(treatment = "NEC",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "FIGURE 7",
                    range = "G5:J10"))))
xxx_wide <- rbind(xxx_wide,
                    data.table(treatment = "D20",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "FIGURE 7",
                    range = "K5:O10"))))
xxx_wide <- rbind(xxx_wide,
                    data.table(treatment = "No cells",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "FIGURE 7",
                    range = "P5:S10"))))
xxx_wide <- rbind(xxx_wide,
                    data.table(treatment = "Human",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "FIGURE 7",
                    range = "T5:V10"))))
setnames(xxx_wide, old=paste0("V", seq_along(times)), new=time_cols)
xxx_wide[, id := factor(1:.N)]
xxx_wide[, treatment := factor(treatment,
                                c("eBCs", "No cells", "NEC", "D20", "Human"))]
Y <- xxx_wide[, .SD, .SDcols = time_cols]
xxx_wide[, auc := apply(Y, 1, auc, x=times)] #check
xxx_wide[, iauc := apply(Y, 1, auc, x=times, method="iauc")]
xxx_wide[, pos.iauc := apply(Y, 1, auc, x=times, method="pos.iauc")]
xxx_wide[, pb.auc := apply(Y, 1, auc, x=times, method="pb.auc")]
xxx_wide[, glucose_mean := apply(Y, 1, mean)]
xxx_wide[, glucose_change := glucose_mean - glucose0]

xxx_long <- melt(xxx_wide, 
                    id.vars = c("treatment", "id"),
                    measure.vars = time_cols,
                    variable.name = "time",
                    value.name = "glucose")
xxx_long[, time:=factor(time)]

```


```{r xxx-n, echo=FALSE, eval=FALSE}
xxx_wide[, .(N=.N), by=.(treatment)]

```

```{r xxx-sigma, echo=FALSE, eval=FALSE}
fit <- gls(glucose ~ time*treatment,
           data = xxx_long,
           weights = varIdent(form= ~ 1 | time),
           correlation= corSymm(form=~ 1| id)
)
round(cov2cor(getVarCov(fit, individual=xxx_long[1,id])), 3)

```

```{r xxx-plot, echo=FALSE, eval=FALSE}
plot_all_curves(xxx_long)
```

```{r xxx-models, echo=FALSE, eval=FALSE}
m0 <- lm(glucose0 ~ treatment, data = xxx_wide)
m1 <- lm(auc ~ treatment, data = xxx_wide)
m2 <- lm(iauc ~ treatment, data = xxx_wide)
m3 <- lm(pos.iauc ~ treatment, data = xxx_wide)
m4 <- lm(pb.auc ~ glucose0 + treatment, data = xxx_wide)
m5 <- lm(glucose_mean ~ glucose0 + treatment, data=xxx_wide)
m6 <- lm(glucose_change ~ treatment, data=xxx_wide)

coef(summary(m0))
coef(summary(m1))
coef(summary(m2))
coef(summary(m3))
coef(summary(m4))
coef(summary(m5))
coef(summary(m6))

```

# Perinatal lead (Pb) exposure results in sex-specific effects on food intake, fat, weight, and insulin response across the murine life-course

1. Sigma
2. sigma relative to initial change from baseline
0.2527439 0.5507195 0.5451951 0.4125610 0.6009756

```{r import-perinatal, message=FALSE}
folder <- "Perinatal lead (Pb) exposure results in sex-specific effects on food intake, fat, weight, and insulin response across the murine life-course"
filename <- "Glucose_Tolerance_Test_CompleteData.xlsx"
file_path <- here(data_path, folder, filename)
lead_wide <- read_excel(file_path,
                    sheet = "Sheet1",
                    range = "A1:Z109") %>%
  clean_names() %>%
  data.table()

times <- c(0, 15, 30, 60, 120)
glucose_cols = paste0("glucose", times)
insulin_cols = paste0("insulin", times)
lead_long <- melt(lead_wide, 
              id.vars = c("id", "sex", "exposure", "glucose_auc", "insulin_auc", "homair_liv"),
              measure.vars = list(glucose_cols, insulin_cols),
              variable.name = "measure_time",
              value.name = c("glucose", "insulin"))
lead_long[, time := times[measure_time]]
setnames(lead_long, old="exposure", new="treatment")
```

```{r perinatal N}
lead_wide[, .(N=.N), by=.(exposure, sex)]
```

```{r perinatal-Sigma}
fit <- gls(glucose ~ factor(time)*sex*treatment,
           data = lead_long,
           weights = varIdent(form= ~ 1 | time),
           correlation= corSymm(form=~ 1| id)
)
mu <- apply(lead_wide[sex == "m" & exposure == "0 ppm", .SD, .SDcols=glucose_cols], 2, mean)
mu
Sigma <- getVarCov(fit, individual=lead_long[1,id])
round(sqrt(diag(Sigma)), 3)
round(sqrt(diag(Sigma)), 3)/(mu[2]-mu[1])
round(cov2cor(Sigma), 3)

```

```{r perinatal-plot}
plot_all_curves(lead_long)
plot_all_curves(lead_long[treatment==unique(treatment)[1]])
plot_all_curves(lead_long[treatment==unique(treatment)[2]])
plot_all_curves(lead_long[treatment==unique(treatment)[3]])

```

# Adrb2 controls glucose homeostasis by developmental regulation of pancreatic islet vasculature

1. Sigma - moderate negative baseline cors, non-baseline cors large positive
1b. sigma relative to initial response
0.06581794 0.39155361 0.25379231 0.28717465
2. treatment has positive effect at T0
3. All tests consistent in direction
4. mediation test has moderate negative indirect effect which is consistent with moderate negative correlations between T0, T1-4

```{r import-ardb2, message=FALSE}
folder <- "Adrb2 controls glucose homeostasis by developmental regulation of pancreatic islet vasculature"
filename <- "elife-39689-fig1-data1-v2.xlsx"
file_path <- here(data_path, folder, filename)

# female controls
dt_wide <- read_excel(file_path, 
                        sheet="Fig1A,B", 
                        range="A6:G9", 
                        col_names = FALSE) %>%
  data.table
n <- ncol(dt_wide) - 1
id_cols <- paste0("cn", 1:n)
colnames(dt_wide) <- c("time", id_cols)
f_cn <- melt(dt_wide,
             id.vars = "time",
             measure.vars = id_cols,
             variable.name = "id",
             value.name = "glucose")

# female treated
dt_wide <- read_excel(file_path, 
                        sheet="Fig1A,B", 
                        range="A12:H15", 
                        col_names = FALSE) %>%
  data.table
n <- ncol(dt_wide) - 1
id_cols <- paste0("tr", 1:n)
colnames(dt_wide) <- c("time", id_cols)
f_tr <- melt(dt_wide,
             id.vars = "time",
             measure.vars = id_cols,
             variable.name = "id",
             value.name = "glucose")

ardb2_long <- rbind(data.table(treatment = "cn", f_cn),
               data.table(treatment = "Ardb2 cKO", f_tr))
ardb2_long[, treatment := factor(treatment, c("cn", "Ardb2 cKO"))]
times <- unique(ardb2_long$time)
time_cols <- paste0("glucose", times)

ardb2_wide <- dcast(ardb2_long, 
                    id + treatment ~ time,
                    value.var = "glucose")
setnames(ardb2_wide, old=as.character(times), new=time_cols)
Y <- ardb2_wide[, .SD, .SDcols=time_cols]
ardb2_wide[, auc:=apply(Y, 1, auc, x=times)]
ardb2_wide[, iauc:=apply(Y, 1, auc, x=times, method="iauc")]
ardb2_wide[, pos.iauc:=apply(Y, 1, auc, x=times, method="pos.iauc")]
ardb2_wide[, glucose_mean := apply(Y, 1, mean)]
ardb2_wide[, glucose_change := glucose_mean - glucose0]

```

```{r ardb2 n}
ardb2_wide[, .(N=.N), by=treatment]
```


```{r ardb2 plot curves}
plot_all_curves(ardb2_long)

```

```{r ardb2 Sigma}
fit <- gls(glucose ~ factor(time)*treatment,
           data = ardb2_long,
           weights = varIdent(form= ~ 1 | time),
           correlation= corSymm(form=~ 1| id)
)
mu <- apply(ardb2_wide[treatment == "cn", .SD, .SDcols=time_cols], 2, mean)
mu
Sigma <- getVarCov(fit, individual=ardb2_long[1,id])
round(sqrt(diag(Sigma)), 3)
round(sqrt(diag(Sigma)), 3)/(mu[2]-mu[1])
round(cov2cor(Sigma), 3)

```

```{r ardb2-models}
m1 <- lm(auc ~ treatment, data = ardb2_wide)
m2 <- lm(iauc ~ treatment, data = ardb2_wide)
m3 <- lm(pos.iauc ~ treatment, data = ardb2_wide)
m4 <- lm(glucose_mean ~ glucose0 + treatment, data=ardb2_wide)
m5 <- lm(glucose_change ~ treatment, data=ardb2_wide)

coef(summary(m1))["treatmentArdb2 cKO",]
coef(summary(m2))["treatmentArdb2 cKO",]
coef(summary(m3))["treatmentArdb2 cKO",]
coef(summary(m4))["treatmentArdb2 cKO",]
coef(summary(m5))["treatmentArdb2 cKO",]

```

```{r ardb2-difference-at-baseline}
m1 <- lm(glucose0 ~ treatment, data=ardb2_wide)
coef(summary(m1))

ggdotplot(data=ardb2_wide,
          x = "treatment",
          y = "glucose0",
          add = "mean_ci")

```

```{r ardb2 direct v indirect}

m1 <- lm(glucose0 ~ treatment, data=ardb2_wide) # mediator model
coef(summary(m1))

m2 <- lm(glucose30 ~ treatment, data=ardb2_wide) # total effect
coef(summary(m2))

m3 <- lm(glucose30 ~ glucose0 + treatment, data=ardb2_wide) # outcome model
coef(summary(m3))

total <- coef(m2)["treatmentArdb2 cKO"]
direct <- coef(m3)["treatmentArdb2 cKO"]
indirect <- total - direct
a <- coef(m1)["treatmentArdb2 cKO"]
b <- coef(m3)["glucose0"]
a
b
a*b
total
direct
indirect

m4 <- mediate(model.m <- m1, 
                   model.y <- m3, 
                   treat = "treatment", 
                   mediator = "glucose0", 
                   robustSE = TRUE, sims = 1000)
summary(m4)

```


# Hepatic Aryl hydrocarbon Receptor Nuclear Translocator (ARNT) regulates metabolism in mice

1. baseline cors very small for gtt, non-baseline cors mod to large positive
1b. sigma relative to initial response in WT
0.12818475 0.08265306 0.30683136 0.12818475 0.08265306 0.30683136
2. AUC, mean-cov suggest positive treatment effect (p=.01, .1) BUT iAUC/mean-change suggest negative treatment effect (p=.44, .48).
3. treatment has positive fasting effect.
4. it looks like there is fasting difference (x0), then about same at x15, x30, then return to fasting difference, so "effect" is not a different response to gtt but simply fasting level...
5. lm-cov model at time15 and time30 show small positive difference (p=.42, .18) not distinguishable from noise

points
1. one cannot blindly analyze AUC but needs to look at data. The AUC interpretation is there is a treatment effect on gtt but the data is more compatible with a fasting effect.

```{r import-arnt-gtt}
folder <- "Hepatic Aryl hydrocarbon Receptor Nuclear Translocator (ARNT) regulates metabolism in mice"

## Fig 1A, D, E
filename <- "Fasting glucose and GTT and AUCGTT.xlsx"
file_path <- here(data_path, folder, filename)
arnt_wide <- read_excel(file_path, 
                        range="A1:J23") %>%
  clean_names() %>%
  data.table()
times <- c(0, 15, 30, 60, 90, 120)
time_cols <- paste0("x", times)
y_cols <- paste0("x", times[times != 0])

arnt_wide[, my_auc := apply(arnt_wide[, .SD, .SDcols=time_cols], 1, auc, x=times)] #check
arnt_wide[, iauc := apply(arnt_wide[, .SD, .SDcols=time_cols], 1, auc, x=times, method="iauc")]
arnt_wide[, ipos.auc := apply(arnt_wide[, .SD, .SDcols=time_cols], 1, auc, x=times, method="pos.iauc")]
arnt_wide[, glucose_mean := apply(arnt_wide[, .SD, .SDcols=y_cols], 1, mean)]
arnt_wide[, glucose_change := glucose_mean - x0]

# there is a miscoded value in cre-wt: "Wt" instead of "WT"
arnt_wide[cre_wt == "Wt", cre_wt := "WT"]
arnt_wide[, treatment := factor(cre_wt, c("WT", "Cre"))]
arnt_wide[, id := as.character(mouse_no)]

arnt_long <- melt(arnt_wide,
                  id.vars = c("id", "treatment", "sex", "auc", "iauc"),
                  measure.vars = time_cols,
                  variable.name = "time",
                  value.name = "glucose")
arnt_long[, time := factor(time)]


```

```{r arnt-n}
arnt_wide[, .(N=.N), by=treatment]
# miscoded "Wt" - go back up and fix
```

```{r arnt-sigma}
fit <- gls(glucose ~ time*treatment,
           data = arnt_long,
           weights = varIdent(form= ~ 1 | time),
           correlation= corSymm(form=~ 1 | id)
)
mu <- apply(arnt_wide[treatment == "WT", .SD, .SDcols=time_cols], 2, mean)
mu
Sigma <- getVarCov(fit, individual=arnt_long[1,id])
round(sqrt(diag(Sigma)), 3)
round(sqrt(diag(Sigma)), 3)/(mu[2]-mu[1])
round(cov2cor(Sigma), 3)

```

```{r arnt-plot-all}
plot_all_curves(arnt_long)
```
 
```{r arnt-models}
m1 <- lm(auc ~ treatment, data = arnt_wide)
m2 <- lm(iauc ~ treatment, data = arnt_wide)
m3 <- lm(ipos.auc ~ treatment, data = arnt_wide)
m4 <- lm(glucose_mean ~ x0 + treatment, data=arnt_wide)
m5 <- lm(glucose_change ~ treatment, data=arnt_wide)

coef(summary(m1))
coef(summary(m2))
coef(summary(m3))
coef(summary(m4))
coef(summary(m5))

```

```{r arnt-difference-at-time0}
m1 <- lm(x0 ~ treatment, data=arnt_wide)
coef(summary(m1))

ggdotplot(data=arnt_wide,
          x = "treatment",
          y = "x0",
          add = "mean_ci")
```

```{r arnt-differences at time15-30-60}
m0 <- lm(x0 ~ treatment, data=arnt_wide)
m1 <- lm(x15 ~ x0 + treatment, data=arnt_wide)
m2 <- lm(x30 ~ x0 + treatment, data=arnt_wide)
m3 <- lm(x60 ~ x0 + treatment, data=arnt_wide)
m4 <- lm(x90 ~ x0 + treatment, data=arnt_wide)
m5 <- lm(x120 ~ x0 + treatment, data=arnt_wide)


m6 <- lm(x15 ~ treatment, data=arnt_wide)
m7 <- lm(x120 ~ treatment, data=arnt_wide)

coef(summary(m0))["treatmentCre",]
coef(summary(m1))["treatmentCre",]
coef(summary(m2))["treatmentCre",]
coef(summary(m3))["treatmentCre",]
coef(summary(m4))["treatmentCre",]
coef(summary(m5))["treatmentCre",]

coef(summary(m6))["treatmentCre",]
coef(summary(m7))["treatmentCre",]

```

```{r import-arnt-ptt}
## fig 1B, C
filename <- "PTT LARNT.xls"
file_path <- here(data_path, folder, filename)
larnt_ptt_wide <- read_excel(file_path, 
                        range="A2:G16") %>%
  data.table()
times_ptt <- c(0, 15, 30, 45, 60)
ptt_times_cols <- paste0("x", times_ptt)
colnames(larnt_ptt_wide) <- c("id", "treatment", ptt_times_cols)
Y <- larnt_ptt_wide[, .SD, .SDcols=ptt_times_cols]
larnt_ptt_wide[, my_auc := apply(Y, 1, auc, x=times_ptt)] #check
larnt_ptt_wide[, iauc := apply(Y, 1, auc, x=times_ptt, method="iauc")]
larnt_ptt_wide[, ipos.auc := apply(Y, 1, auc, x=times_ptt, method="pos.iauc")]
larnt_ptt_wide[, glucose_mean := apply(Y, 1, mean)]
larnt_ptt_wide[, treatment := factor(treatment)]

larnt_ptt_long <- melt(larnt_ptt_wide,
                  id.vars = c("id", "treatment", "my_auc", "iauc"),
                  measure.vars = ptt_times_cols <- paste0("x", times_ptt),
                  variable.name = "time",
                  value.name = "glucose")

```

```{r ptt-models}
m1 <- lm(my_auc ~ treatment, data = larnt_ptt_wide)
m2 <- lm(iauc ~ treatment, data = larnt_ptt_wide)
m3 <- lm(ipos.auc ~ treatment, data = larnt_ptt_wide)
m4 <- lm(glucose_mean ~ x0 + treatment, data=larnt_ptt_wide)

coef(summary(m1))
coef(summary(m2))
coef(summary(m3))
coef(summary(m4))

t.test(my_auc ~ treatment, data = larnt_ptt_wide, var.equal=FALSE)
# my lm/t.test differs because the authors computed a
# different AUC for mouse 106 (same for all others)
# regardles mouse 106 time 1 measure is an extreme outlier
# re-run without this

m1 <- lm(my_auc ~ treatment, data = larnt_ptt_wide[id != 106])
m4 <- lm(glucose_mean ~ x0 + treatment, data=larnt_ptt_wide[id != 106])
coef(summary(m1))
coef(summary(m4))

```

Authors computed something like a positive incremental area but not correctly.

```{r larnt-ptt-plot}
plot_all_curves(larnt_ptt_long)
```

# Suppression of insulin feedback enhances the efficacy of PI3K inhibitors

1. small correlations in Sigma. Noise is apparent from plot
1b. sd relative to initial response in control
0.4574595 0.8412703 1.6563243 2.3108919 1.9452703 2.0679189 1.6714054 1.7787838
2. treatment effect consistent

```{r import-pi3k}
folder <- "Suppression of insulin feedback enhances the efficacy of PI3K inhibitors"
filename <- "41586_2018_343_MOESM4_ESM.xlsx"
file_path <- here(data_path, folder, filename)
pi3k_wide <- read_excel(file_path,
                    col_names = FALSE,
                    sheet = "Figure 1A (Glucose Curves)",
                    range = "A2:U9") %>%
  data.table()

# give columns names
colnames(pi3k_wide) <- c(
  "time",
  paste0("Vehicle_", 1:5),
  paste0("BKM120", 1:5),
  paste0("BYL719", 1:5),
  paste0("GDC0032", 1:5)
)

# melt 1
pi3k_long <- melt(pi3k_wide, 
              id.vars = c("time"),
              measure.vars = patterns("^Vehicle",
                                      "^BKM120",
                                      "^BYL719",
                                      "GDC0032"),
              variable.name = "rep",
              value.name = c("Vehicle",
                                "BKM120",
                                "BYL719",
                                "GDC0032"))

# melt 2
pi3k_long <- melt(pi3k_long, 
              id.vars = c("time", "rep"),
              measure.vars = c("Vehicle",
                                "BKM120",
                                "BYL719",
                                "GDC0032"),
              variable.name = "treatment",
              value.name = "glucose")
pi3k_long[, time:=factor(time)]
pi3k_long[, id := paste(treatment, rep, sep="_")]

pi3k_wide <- dcast(data = pi3k_long,
                    id + treatment + rep ~ time, value.var = "glucose")
times <- c(0, 15, 30, 60, 90, 120, 180, 240)
time_cols <- paste0("glucose", times)
setnames(pi3k_wide, old=as.character(times), new=time_cols)
Y <- pi3k_wide[, .SD, .SDcols = time_cols]
pi3k_wide[, auc := apply(Y, 1, auc, x=times)] #check
pi3k_wide[, iauc := apply(Y, 1, auc, x=times, method="iauc")]
pi3k_wide[, ipos.auc := apply(Y, 1, auc, x=times, method="pos.iauc")]
pi3k_wide[, glucose_mean := apply(Y, 1, mean)]
pi3k_wide[, glucose_change := glucose_mean - glucose0]

```


```{r pi3k-n}
pi3k_wide[, .(N=.N), by=.(treatment)]

```

```{r pi3k-sigma}
fit <- gls(glucose ~ time*treatment,
           data = pi3k_long,
           weights = varIdent(form= ~ 1 | time),
           correlation= corSymm(form=~ 1| id)
)
mu <- apply(pi3k_wide[treatment == "Vehicle", .SD, .SDcols=time_cols], 2, mean)
mu
Sigma <- getVarCov(fit, individual=pi3k_long[1,id])
round(sqrt(diag(Sigma)), 3)
round(sqrt(diag(Sigma)), 3)/(mu[2]-mu[1])
round(cov2cor(Sigma), 3)

```

```{r pi3k-plot}
plot_all_curves(pi3k_long)
```

```{r pi3k-models}
m1 <- lm(auc ~ treatment, data = pi3k_wide)
m2 <- lm(iauc ~ treatment, data = pi3k_wide)
m3 <- lm(ipos.auc ~ treatment, data = pi3k_wide)
m4 <- lm(glucose_mean ~ glucose0 + treatment, data=pi3k_wide)
m5 <- lm(glucose_change ~ treatment, data=pi3k_wide)

coef(summary(m1))
coef(summary(m2))
coef(summary(m3))
coef(summary(m4))
coef(summary(m5))

```

# The E3 ubiquitin ligase MARCH1 regulates glucose-tolerance and lipid storage in a sex-specific manner

1. very small baseline cor, mod to mod-large non-baseline cors
1b. sigma relative to initial response in WT
0.09597674 0.31565550 0.18280642 0.27442647 0.24691554

```{r import-March1}
folder <- "The E3 ubiquitin ligase MARCH1 regulates glucose-tolerance and lipid storage in a sex-specific manner"
filename <- "Figure2_data.xlsx"
file_path <- here(data_path, folder, filename)
times <- c(0, 20, 40, 60, 90)
time_cols <- paste0("glucose", times)

march1_wide <- data.table(sex = "male",
                       treatment = "wt",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "Panels A and B",
                    range = "B2:I6")))
march1_wide <- rbind(march1_wide,
                    data.table(sex = "male",
                       treatment = "march1-ko",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "Panels A and B",
                    range = "j2:p6"))))
march1_wide <- rbind(march1_wide,
                    data.table(sex = "female",
                       treatment = "wt",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "Panels A and B",
                    range = "R2:AA6"))))
march1_wide <- rbind(march1_wide,
                    data.table(sex = "female",
                       treatment = "march1-ko",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "Panels A and B",
                    range = "AB2:AM6"))))
setnames(march1_wide, old=paste0("V", 1:5), new=time_cols)
march1_wide[, id := factor(as.character(1:.N))]
march1_wide[, treatment := factor(treatment)]
march1_wide[, sex := factor(sex)]

Y <- march1_wide[, .SD, .SDcols = time_cols]
march1_wide[, auc := apply(Y, 1, auc, x=times)] #check
march1_wide[, iauc := apply(Y, 1, auc, x=times, method="iauc")]
march1_wide[, ipos.auc := apply(Y, 1, auc, x=times, method="pos.iauc")]
march1_wide[, glucose_mean := apply(Y, 1, mean)]
march1_wide[, glucose_change := glucose_mean - glucose0]

march1_long <- melt(march1_wide, 
                    id.vars = c("sex", "treatment", "id"),
                    measure.vars = time_cols,
                    variable.name = "time",
                    value.name = "glucose")
march1_long[, time := factor(time)]

```


```{r march1-n}
march1_wide[, .(N=.N), by=.(treatment, sex)]

```

```{r march1-sigma}
fit <- gls(glucose ~ sex*time*treatment,
           data = march1_long,
           weights = varIdent(form= ~ 1 | time),
           correlation= corSymm(form=~ 1| id)
)
mu <- apply(march1_wide[treatment == "wt", .SD, .SDcols=time_cols], 2, mean)
mu
Sigma <- getVarCov(fit, individual=march1_long[1,id])
round(sqrt(diag(Sigma)), 3)
round(sqrt(diag(Sigma)), 3)/(mu[2]-mu[1])
round(cov2cor(Sigma), 3)

```

```{r march1-plot}
plot_all_curves(march1_long)
```

```{r march-models}
m1 <- lm(auc ~ treatment, data = march1_wide[sex=="male"])
m2 <- lm(iauc ~ treatment, data = march1_wide[sex=="male"])
m3 <- lm(ipos.auc ~ treatment, data = march1_wide[sex=="male"])
m4 <- lm(glucose_mean ~ glucose0 + treatment, data=march1_wide[sex=="male"])
m5 <- lm(glucose_change ~ treatment, data=march1_wide[sex=="male"])
subdata <- march1_long[sex == "male"]
m6 <- aov_4(glucose ~ time*treatment + (time|id),
                 data = subdata)
m7 <- aov_4(glucose ~ time*treatment + (time|id),
                 data = subdata, anova_table = list(correction = "none"))
m8 <- gls(glucose ~ treatment*time,
           weights = varIdent(form= ~ 1 | time),
          correlation= corCompSymm(form = ~ 1|id),
          data=subdata)
m9 <- lmer(glucose ~ treatment*time + (1|id),
          data=subdata)
m6$anova_table
m7$anova_table
anova(m8)
anova(m9)
emmeans(m7, revpairwise ~ treatment | time)
contrast(emmeans(m7, specs = ))


coef(summary(m1))
coef(summary(m2))
coef(summary(m3))
coef(summary(m4))
coef(summary(m5))

```

# Endogenous nicotinamide riboside metabolism protects against diet-induced liver damage

ptt fig 1
1. baseline cors are mod to mod-large positive. glucose15 cor is mod positive with glucose0 and glucose30 but very low with everything else. Non-baseline cors after glucose15 mod to large positive
2. Treatment effects very consistent among models except that auc has high SE and so p = .09 while other p = .02

gtt fig 2
1. mistake in pval for iuc for gtt. The auc are incorrectly assigned to treatment -- see excel file and compare to my results. Good example of excel problems.
2. baseline cor high pos to low from t15 to t180, non-baseline cor have pattern in fake_rho simulation (high to mod) except t15 is mod to very low.

## nrk1 ptt
```{r import-nrk1}
folder <- "Endogenous nicotinamide riboside metabolism protects against diet-induced liver damage"
filename <- "41467_2019_12262_MOESM4_ESM.xlsx"
file_path <- here(data_path, folder, filename)
times <- c(0, 15, 30, 45, 60, 90, 120)
time_cols <- paste0("glucose", times)

nrk1_wide <- data.table(treatment = "wt",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "Figure 1",
                    range = "C4:H10")))
nrk1_wide <- rbind(nrk1_wide,
                    data.table(treatment = "nrk1-ko",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "Figure 1",
                    range = "I4:O10"))))
setnames(nrk1_wide, old=paste0("V", seq_along(times)), new=time_cols)
nrk1_wide[, id := as.character(1:.N)]
nrk1_wide[, treatment := factor(treatment, c("wt", "nrk1-ko"))]
Y <- nrk1_wide[, .SD, .SDcols = time_cols]
nrk1_wide[, auc := apply(Y, 1, auc, x=times)] #check
nrk1_wide[, iauc := apply(Y, 1, auc, x=times, method="iauc")]
nrk1_wide[, pos.iauc := apply(Y, 1, auc, x=times, method="pos.iauc")]
nrk1_wide[, pb.auc := apply(Y, 1, auc, x=times, method="pb.auc")]
nrk1_wide[, glucose_mean := apply(Y, 1, mean)]
nrk1_wide[, glucose_change := glucose_mean - glucose0]

nrk1_long <- melt(nrk1_wide, 
                    id.vars = c("treatment", "id"),
                    measure.vars = time_cols,
                    variable.name = "time",
                    value.name = "glucose")
```


```{r nrk1-n}
nrk1_wide[, .(N=.N), by=.(treatment)]

```

```{r nrk1-sigma}
fit <- gls(glucose ~ time*treatment,
           data = nrk1_long,
           weights = varIdent(form= ~ 1 | time),
           correlation= corSymm(form=~ 1| id)
)
round(cov2cor(getVarCov(fit, individual=nrk1_long[1,id])), 3)

```

```{r nrk1-plot}
plot_all_curves(nrk1_long)
```

```{r nrk1-models}
m0 <- lm(glucose0 ~ treatment, data = nrk1_wide)
m1 <- lm(auc ~ treatment, data = nrk1_wide)
m2 <- lm(iauc ~ treatment, data = nrk1_wide)
m3 <- lm(pos.iauc ~ treatment, data = nrk1_wide)
m4 <- lm(pb.auc ~ glucose0 + treatment, data = nrk1_wide)
m5 <- lm(glucose_mean ~ glucose0 + treatment, data=nrk1_wide)
m6 <- lm(glucose_change ~ treatment, data=nrk1_wide)

coef(summary(m0))["treatmentnrk1-ko",]
coef(summary(m1))["treatmentnrk1-ko",]
coef(summary(m2))["treatmentnrk1-ko",]
coef(summary(m3))["treatmentnrk1-ko",]
coef(summary(m4))["treatmentnrk1-ko",]
coef(summary(m5))["treatmentnrk1-ko",]
coef(summary(m6))["treatmentnrk1-ko",]

```

## nrk1 gtt


```{r import-nrk1}
folder <- "Endogenous nicotinamide riboside metabolism protects against diet-induced liver damage"
filename <- "41467_2019_12262_MOESM4_ESM.xlsx"
file_path <- here(data_path, folder, filename)
times <- c(0, 15, 30, 45, 60, 90, 120, 150, 180)
time_cols <- paste0("glucose", times)

nrk1_wide <- data.table(treatment = "cn",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "Figure 2",
                    range = "D66:K74")))
nrk1_wide <- rbind(nrk1_wide,
                    data.table(treatment = "nrk1-ko",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "Figure 2",
                    range = "L66:U74"))))
setnames(nrk1_wide, old=paste0("V", seq_along(times)), new=time_cols)
nrk1_wide[, id := as.character(1:.N)]
nrk1_wide[, treatment := factor(treatment, c("cn", "nrk1-ko"))]
Y <- nrk1_wide[, .SD, .SDcols = time_cols]
nrk1_wide[, auc := apply(Y, 1, auc, x=times)] #check
nrk1_wide[, iauc := apply(Y, 1, auc, x=times, method="iauc")]
nrk1_wide[, pos.iauc := apply(Y, 1, auc, x=times, method="pos.iauc")]
nrk1_wide[, pb.auc := apply(Y, 1, auc, x=times, method="pb.auc")]
nrk1_wide[, glucose_mean := apply(Y, 1, mean)]
nrk1_wide[, glucose_change := glucose_mean - glucose0]

nrk1_long <- melt(nrk1_wide, 
                    id.vars = c("treatment", "id"),
                    measure.vars = time_cols,
                    variable.name = "time",
                    value.name = "glucose")
```


```{r nrk1-n}
nrk1_wide[, .(N=.N), by=.(treatment)]

```

```{r nrk1-sigma}
fit <- gls(glucose ~ time*treatment,
           data = nrk1_long,
           weights = varIdent(form= ~ 1 | time),
           correlation= corSymm(form=~ 1| id)
)
mu <- apply(nrk1_wide[treatment == "cn", .SD, .SDcols=time_cols], 2, mean)
mu
Sigma <- getVarCov(fit, individual=nrk1_long[1,id])
round(sqrt(diag(Sigma)), 3)
round(sqrt(diag(Sigma)), 3)/(mu[2]-mu[1])
round(cov2cor(Sigma), 3)

```

```{r nrk1-plot}
plot_all_curves(nrk1_long)
```

```{r nrk1-models}
m0 <- lm(glucose0 ~ treatment, data = nrk1_wide)
m1 <- lm(auc ~ treatment, data = nrk1_wide)
m2 <- lm(iauc ~ treatment, data = nrk1_wide)
m3 <- lm(pos.iauc ~ treatment, data = nrk1_wide)
m4 <- lm(pb.auc ~ glucose0 + treatment, data = nrk1_wide)
m5 <- lm(glucose_mean ~ glucose0 + treatment, data=nrk1_wide)
m6 <- lm(glucose_change ~ treatment, data=nrk1_wide)

coef(summary(m0))["treatmentnrk1-ko",]
coef(summary(m1))["treatmentnrk1-ko",]
coef(summary(m2))["treatmentnrk1-ko",]
coef(summary(m3))["treatmentnrk1-ko",]
coef(summary(m4))["treatmentnrk1-ko",]
coef(summary(m5))["treatmentnrk1-ko",]
coef(summary(m6))["treatmentnrk1-ko",]

```

# Recapitulating endocrine cell clustering in culture promotes maturation of human stem-cell-derived β cells
1. Very small baseline cor and mod to large non baseline cor
1b. sigma relative to initial response in "no cells"
0.04268047 0.18612678 0.04268047 0.18612678 0.04268047 0.18612678
2. heterogenity of baseline glucose, eith eBCs very low

```{r import-beta}
folder <- "Recapitulating endocrine cell clustering in culture promotes maturation of human stem-cell-derived β cells"
filename <- "41556_2018_271_MOESM8_ESM.xlsx"
file_path <- here(data_path, folder, filename)
times <- c(0, 15, 30, 60, 90, 120)
time_cols <- paste0("glucose", times)

beta_wide <- data.table(treatment = "eBCs",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "FIGURE 7",
                    range = "B5:F10")))
beta_wide <- rbind(beta_wide,
                    data.table(treatment = "NEC",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "FIGURE 7",
                    range = "G5:J10"))))
beta_wide <- rbind(beta_wide,
                    data.table(treatment = "D20",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "FIGURE 7",
                    range = "K5:O10"))))
beta_wide <- rbind(beta_wide,
                    data.table(treatment = "No cells",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "FIGURE 7",
                    range = "P5:S10"))))
beta_wide <- rbind(beta_wide,
                    data.table(treatment = "Human",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "FIGURE 7",
                    range = "T5:V10"))))
setnames(beta_wide, old=paste0("V", seq_along(times)), new=time_cols)
beta_wide[, id := factor(1:.N)]
beta_wide[, treatment := factor(treatment,
                                c("eBCs", "No cells", "NEC", "D20", "Human"))]
Y <- beta_wide[, .SD, .SDcols = time_cols]
beta_wide[, auc := apply(Y, 1, auc, x=times)] #check
beta_wide[, iauc := apply(Y, 1, auc, x=times, method="iauc")]
beta_wide[, pos.iauc := apply(Y, 1, auc, x=times, method="pos.iauc")]
beta_wide[, pb.auc := apply(Y, 1, auc, x=times, method="pb.auc")]
beta_wide[, glucose_mean := apply(Y, 1, mean)]
beta_wide[, glucose_change := glucose_mean - glucose0]

beta_long <- melt(beta_wide, 
                    id.vars = c("treatment", "id"),
                    measure.vars = time_cols,
                    variable.name = "time",
                    value.name = "glucose")
```


```{r beta-n}
beta_wide[, .(N=.N), by=.(treatment)]

```

```{r beta-sigma}
fit <- gls(glucose ~ time*treatment,
           data = beta_long,
           weights = varIdent(form= ~ 1 | time),
           correlation= corSymm(form=~ 1| id)
)
mu <- apply(beta_wide[treatment == "No cells", .SD, .SDcols=time_cols], 2, mean)
mu
Sigma <- getVarCov(fit, individual=beta_long[1,id])
round(sqrt(diag(Sigma)), 3)
round(sqrt(diag(Sigma)), 3)/(mu[2]-mu[1])
round(cov2cor(Sigma), 3)

```

```{r beta-plot}
plot_all_curves(beta_long)
```

```{r beta-models}
m0 <- lm(glucose0 ~ treatment, data = beta_wide)
m1 <- lm(auc ~ treatment, data = beta_wide)
m2 <- lm(iauc ~ treatment, data = beta_wide)
m3 <- lm(pos.iauc ~ treatment, data = beta_wide)
m4 <- lm(pb.auc ~ glucose0 + treatment, data = beta_wide)
m5 <- lm(glucose_mean ~ glucose0 + treatment, data=beta_wide)
m6 <- lm(glucose_change ~ treatment, data=beta_wide)

coef(summary(m0))
coef(summary(m1))
coef(summary(m2))
coef(summary(m3))
coef(summary(m4))
coef(summary(m5))
coef(summary(m6))

```

# BCAA catabolism in brown fat controls energy homeostasis through SLC25A44

1. moderate to moderately high baseline cor and moderate-very high non-baseline cor
1b. sigma relative to initial response in control
0.1276123 0.2923751 0.2841936 0.1276123 0.2923751 0.2841936
2. some evidence of increased glucose at baseline in treatment (p ~ 0.05)
3. evidence of glucose intolerance less compelling using covariate model than not using covariate model (p ~ 0.05 v .005)
4. Good candidate for difference in estimate based on cov v. non cov model. Mediation model shows >50% of effect is indirect
5. Good example of regression to the mean at glucose15 when there is a baseline difference (is this real?)

```{r import-bcaa}
folder <- "BCAA catabolism in brown fat controls energy homeostasis through SLC25A44"
filename <- "41586_2019_1503_MOESM8_ESM.xlsx"
file_path <- here(data_path, folder, filename)
times <- c(0, 15, 30, 60, 90, 120)
time_cols <- paste0("glucose", times)

bcaa_wide <- read_excel(file_path,
                        sheet = "Fig2g",
                        range = "A2:G30") %>%
  data.table()
colnames(bcaa_wide) <- c("id", time_cols)
bcaa_wide[, id := factor(id)]
bcaa_wide[, treatment := factor(ifelse(substr(id,1,1)=="C", "Control", "BckdhaUcp1 KO"),
                                c("Control", "BckdhaUcp1 KO"))]

Y <- bcaa_wide[, .SD, .SDcols = time_cols]
bcaa_wide[, auc := apply(Y, 1, auc, x=times)] #check
bcaa_wide[, iauc := apply(Y, 1, auc, x=times, method="iauc")]
bcaa_wide[, pos.iauc := apply(Y, 1, auc, x=times, method="pos.iauc")]
bcaa_wide[, pb.auc := apply(Y, 1, auc, x=times, method="pb.auc")]
bcaa_wide[, glucose_mean := apply(Y, 1, mean)]
bcaa_wide[, glucose_change := glucose_mean - glucose0]

bcaa_long <- melt(bcaa_wide, 
                    id.vars = c("treatment", "id", "glucose0"),
                    measure.vars = time_cols,
                    variable.name = "time",
                    value.name = "glucose")
bcaa_long[, time := factor(time)]
bcaa_long[, minutes := times[as.integer(time)]]
```


```{r bcaa-n}
bcaa_wide[, .(N=.N), by=.(treatment)]

```

```{r bcaa-sigma}
fit <- gls(glucose ~ time*treatment,
           data = bcaa_long,
           weights = varIdent(form= ~ 1 | time),
           correlation= corSymm(form=~ 1| id)
)
mu <- apply(bcaa_wide[treatment == "Control", .SD, .SDcols=time_cols], 2, mean)
mu
Sigma <- getVarCov(fit, individual=bcaa_long[1,id])
round(sqrt(diag(Sigma)), 3)
round(sqrt(diag(Sigma)), 3)/(mu[2]-mu[1])
round(cov2cor(Sigma), 3)

```

```{r bcaa-plot}
plot_all_curves(bcaa_long)
```

```{r bcaa-models}
m0 <- lm(glucose0 ~ treatment, data = bcaa_wide)
m1 <- lm(auc ~ treatment, data = bcaa_wide)
m2 <- lm(iauc ~ treatment, data = bcaa_wide)
m3 <- lm(pos.iauc ~ treatment, data = bcaa_wide)
m4 <- lm(pb.auc ~ glucose0 + treatment, data = bcaa_wide)
m5 <- lm(glucose_mean ~ glucose0 + treatment, data=bcaa_wide)
m6 <- lm(glucose_change ~ treatment, data=bcaa_wide)

coef(summary(m0))
coef(summary(m1))
coef(summary(m2))
coef(summary(m3))
coef(summary(m4))
coef(summary(m5))
coef(summary(m6))

```

```{r bcaa-rmanova}
m1 <- aov_4(glucose ~ time*treatment + (time|id),
                 data = bcaa_long)
m2 <- aov_4(glucose ~ time*treatment + (time|id),
                 data = bcaa_long, anova_table = list(correction = "none"))

m1$anova_table
m2$anova_table
        m1.emm <- emmeans(m1,  ~ treatment*time)
        contrast(m1.emm, 
                 method = "revpairwise",
                 simple = "each",
                 combine = TRUE,
                 adjust = "none")[1:6,]
```

```{r bcaa-Gls-setup}
subdata <- bcaa_long[time != "glucose0"]
cp <- list(corCAR1,corExp,corCompSymm,corLin,corGaus,corSpher)
z <- vector( 'list', length(cp))
for(k in 1:length(cp)){
  z[[k]] <- gls(glucose ~ treatment * rcs(minutes, 3) +
                  rcs(glucose0, 3), data=subdata,
                correlation =cp[[k]](form = ~minutes | id))
}
anova(z[[1]],z[[2]],z[[3]],z[[4]],z[[5]],z[[6]])

```

```{r bcaa-Gls}

ddname <- datadist(subdata)
options(datadist="ddname")

m1 <- Gls(glucose ~ treatment * rcs(minutes, 3) +
                  rcs(glucose0, 3), data=subdata,
                correlation =corExp(form = ~minutes | id))
m1
```

```{r bcaa-Gls-pairs}

(m1.pairs <- rms::contrast(m1, 
               list(minutes=c(15, 30, 60, 90, 120), treatment = "Control"),
               list(minutes=c(15, 30, 60, 90, 120), treatment = "BckdhaUcp1 KO")))

```

glucose0   .         BckdhaUcp1 KO - Control     20.7 21.5 50.2 0.964   0.3397 
 glucose15  .         BckdhaUcp1 KO - Control     59.7 21.5 50.2 2.773   0.0078 
 glucose30  .         BckdhaUcp1 KO - Control     67.5 21.5 50.2 3.136   0.0029 
 glucose60  .         BckdhaUcp1 KO - Control     63.5 21.5 50.2 2.951   0.0048 
 glucose90  .         BckdhaUcp1 KO - Control     69.6 21.5 50.2 3.234   0.0022 
 glucose120 .         BckdhaUcp1 KO - Control     48.7 21.5 50.2 2.266   0.0278 

```{r bcaa-mediation-15}
m1 <- lm(glucose0 ~ treatment, data=bcaa_wide) # mediator model

m2 <- lm(glucose15 ~ treatment, data=bcaa_wide) # total effect

m3 <- lm(glucose15 ~ glucose0 + treatment, data=bcaa_wide) # outcome model

m4 <- mediate(model.m <- m1, 
                   model.y <- m3, 
                   treat = "treatment", 
                   mediator = "glucose0", 
                   robustSE = TRUE, sims = 1000)
summary(m4)

```

```{r bcaa-mediation-mean}
m1 <- lm(glucose0 ~ treatment, data=bcaa_wide) # mediator model

m2 <- lm(glucose_mean ~ treatment, data=bcaa_wide) # total effect

m3 <- lm(glucose_mean ~ glucose0 + treatment, data=bcaa_wide) # outcome model

m4 <- mediate(model.m <- m1, 
                   model.y <- m3, 
                   treat = "treatment", 
                   mediator = "glucose0", 
                   robustSE = TRUE, sims = 1000)
summary(m4)


```

# Interleukin-36 cytokines alter the intestinal microbiome and can protect against obesity and metabolic dysfunction

1. baseline cor has small to large *negative* cor. non-baseline cor has small (glucose15) to large (glucose30+) positive cor.

```{r import-il36}
folder <- "Interleukin-36 cytokines alter the intestinal microbiome and can protect against obesity and metabolic dysfunction"
filename <- "41467_2019_11944_MOESM3_ESM.xlsx"
file_path <- here(data_path, folder, filename)
times <- c(0, 15, 45, 70, 120)
time_cols <- paste0("glucose", times)

il36_wide <- data.table(treatment = "wt",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "Figure 2d",
                    range = "C5:G9")))
il36_wide <- rbind(il36_wide,
                    data.table(treatment = "il36rn-/-",
                       t(read_excel(file_path,
                    col_names = FALSE,
                    sheet = "Figure 2d",
                    range = "H5:L9"))))

setnames(il36_wide, old=paste0("V", seq_along(times)), new=time_cols)
il36_wide[, id := factor(1:.N)]
il36_wide[, treatment := factor(treatment,
                                c("wt", "il36rn-/-"))]
Y <- il36_wide[, .SD, .SDcols = time_cols]
il36_wide[, auc := apply(Y, 1, auc, x=times)] #check
il36_wide[, iauc := apply(Y, 1, auc, x=times, method="iauc")]
il36_wide[, pos.iauc := apply(Y, 1, auc, x=times, method="pos.iauc")]
il36_wide[, pb.auc := apply(Y, 1, auc, x=times, method="pb.auc")]
il36_wide[, glucose_mean := apply(Y, 1, mean)]
il36_wide[, glucose_change := glucose_mean - glucose0]

il36_long <- melt(il36_wide, 
                    id.vars = c("treatment", "id"),
                    measure.vars = time_cols,
                    variable.name = "time",
                    value.name = "glucose")
il36_long[, time:=factor(time)]
```


```{r il36-n}
il36_wide[, .(N=.N), by=.(treatment)]

```

```{r il36-sigma}
fit <- gls(glucose ~ time*treatment,
           data = il36_long,
           weights = varIdent(form= ~ 1 | time),
           correlation= corSymm(form=~ 1| id)
)
mu <- apply(il36_wide[treatment == "wt", .SD, .SDcols=time_cols], 2, mean)
mu
Sigma <- getVarCov(fit, individual=il36_long[1,id])
round(sqrt(diag(Sigma)), 3)
round(sqrt(diag(Sigma)), 3)/(mu[2]-mu[1])
round(cov2cor(Sigma), 3)

```

```{r il36-plot}
plot_all_curves(il36_long)
```

```{r il36-models}
m0 <- lm(glucose0 ~ treatment, data = il36_wide)
m1 <- lm(auc ~ treatment, data = il36_wide)
m2 <- lm(iauc ~ treatment, data = il36_wide)
m3 <- lm(pos.iauc ~ treatment, data = il36_wide)
m4 <- lm(pb.auc ~ glucose0 + treatment, data = il36_wide)
m5 <- lm(glucose_mean ~ glucose0 + treatment, data=il36_wide)
m6 <- lm(glucose_change ~ treatment, data=il36_wide)

coef(summary(m0))
coef(summary(m1))
coef(summary(m2))
coef(summary(m3))
coef(summary(m4))
coef(summary(m5))
coef(summary(m6))

```

```{r il36-rmanova}
m1 <- aov_4(glucose ~ time*treatment + (time|id),
                 data = il36_long)
m2 <- aov_4(glucose ~ time*treatment + (time|id),
                 data = il36_long, anova_table = list(correction = "none"))

m1$anova_table
m2$anova_table
        m1.emm <- emmeans(m1,  ~ treatment*time)
        contrast(m1.emm, 
                 method = "revpairwise",
                 simple = "each",
                 combine = TRUE,
                 adjust = "none")[1:6,]
```

